---
title: "Methods to estimate cholera infection and attack rates"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: true
    df_print: paged
---

```{r echo = FALSE, warning = FALSE, include = FALSE}
# Setup Environment
knitr::opts_chunk$set(
  echo = TRUE,
  include = TRUE,
  eval = TRUE,
  warning = TRUE,
  message = FALSE,
  cache = F,
  prompt = F,
  comment = NA,
  fig.width = 7,
  fig.height = 4
  )

# Use english date format even with french or german locale.
Sys.setenv(LANGUAGE="en")
Sys.setlocale("LC_TIME", "en_GB.UTF-8")
```

# Goal

Explore different methods to compute the attack rate of cholera from the serological and surveillance data from [Jackson et al. (2013)](https://doi.org/10.4269%2Fajtmh.13-0208).

This includes methods to estimate the
    - the clinical attack rate of cholera
    - infection rate of cholera (i.e. including mild and asymptomatic infections)

# Libraries and functions

```{r}
# install.packages(c("here", "dplyr", "tidyr", "ggplot2", "magrittr", "rstan", "readr"))

library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(rstan)

source(here("src/R/functions.R"))
```

# Data

```{r}
raw <- readr::read_csv(here("data/serology_clean.csv"))

levels_raw_data <- c(1.414,10*2^seq(11),40560) #titer levels as given in data. 1.414 is an assigned value for lowest dilution. 40560 is an error and should be 40960

#shape data for analysis
data <- raw %>%
  transmute(
    age,
    vibriocidalMAX_titer = factor(vibriocidalMAX_titer, levels = levels_raw_data, ordered = TRUE),
    n
  ) %>%
  uncount(n)


#assign 1 to titers not detectable
#change wrong label for highest dilution
levels(data$vibriocidalMAX_titer)[levels(data$vibriocidalMAX_titer)=="40560"] <- "40960"
levels(data$vibriocidalMAX_titer)[levels(data$vibriocidalMAX_titer)=="1.414"] <- "1"

#transform titers to log2 scale
data %<>%
  mutate(
    vibriocidalMAX_titer_log2 = log2(as.integer(as.character(vibriocidalMAX_titer)))
  )
```

# Attack rate estimates

## Clinical attack rate

- Total reported cases in Grande Saline commune between the start of the outbreak and 6 April 2011, divided by population.
- Stratification by age group.
- Cholera cases reported to the national surveillance system, as in Figure 2B of [Jackson et al. (2013)](https://doi.org/10.4269%2Fajtmh.13-0208).
- Population from [IHSI 2009](https://www.humanitarianresponse.info/en/operations/haiti/document/projection-de-la-population-par-la-section-communale-de-2009-ihsi).
- Age breakdown from the 2010 Haiti Demographics from [Demography of Haiti](https://en.wikipedia.org/wiki/Demographics_of_Haiti), based on IHSI, retreived on 10.03.2023

```{r}
cases <- 3880
cases_u5 <- 848
cases_2to4 <- round(cases_u5*0.7394) #adjusting for age distribution of cases according to patients in Gonaives CTC Nov-Dec 2010

population <- 21131
population_u5 <- round(population * 12.53/100)
population_2to4 <- round(population_u5*3/5) #rough approximation given regular age distribution in under 5s

attackRate_clinical <- cases/population
attackRate_clinical_u5 <- cases_u5/population_u5
attackRate_clinical_2to4 <- cases_2to4/population_2to4
attackRate_clinical_o5 <- (cases - cases_u5)/(population - population_u5)
attackRate_clinical_o2 <- (cases - (cases_u5-cases_2to4))/(population - (population_u5-population_2to4))
```



## Attack rate from self-reported episodes of cholera

As reported in survey by [Jackson et al. (2013)](https://doi.org/10.4269%2Fajtmh.13-0208).

- `chol` stands for self reported cholera
- `wd` stands for self reported watery diarrhea

```{r}
ar_self_rep <- list()

ar_self_rep$chol <- 466/(2074 + 466 + 82)
ar_self_rep$wd <- 541/(1998 + 541 + 83)
ar_self_rep$chol_2to4 <- 23/(97 + 23 + 4)
ar_self_rep$wd_2to4 <- 31/(89 + 31 + 4)
ar_self_rep$chol_geq5 <- 443/(1974 + 443 + 78)
ar_self_rep$wd_geq5 <- 510/(1906 + 510 + 79)
```


## Threshold on vibriocidal titer

In [Azman et al.](http://doi.org/10.1126/scitranslmed.aau6242) the authors report the best performing threshold to identify recent cholera infection from different
serological markers, alongside with the sensitivity and specificity obtained by using this threshold.

The best threshold depdens on the time period of interest within which cholera infection should be detected. Here we will use the thresholds for detecting
cholera infection within the last 200 days, which corresponds best to our period of interest (about 6 months from the start of the outbreak to the serosurvey).

Two datasets are analyzed, one of 1569 samples of a cohort in Bangladesh, and one from 38 healthy North American volunteers.
Whereas the Bangladesh cohort has a much larger sample size, the high endemicity of cholera in Bangladesh could lead to biased results when applied to the case of Haiti,
where the population was cholera naive before the outbreak.

Using the data from North American volunteers, who are also cholera naive may thus be more apropriate.

We apply the threshold of 320 used to identify recent cholera infection during the last 200 days in [Azman et al.](http://doi.org/10.1126/scitranslmed.aau6242) and adjust for the reported sensitivity and specificity,
using values derived from the dataset of the Bangladesh cohort and the North American Volunteers.

We use the [Rogan-Gladen estimator](https://doi.org/10.1093/oxfordjournals.aje.a112510) (function `correctForSensSpec()`) to adjust for sensitivity and specificity and truncate the result to the interval [0,1]. Uncertainty on sensitivity and
specificity is propagated through the Rogan-Gladen estimator and the min and max values obtained is reported.

### For the entire population

```{r}
thresh <- 320 #corresponds to best threshold to identify cholera during the last 200 days according to table S3 Azman et al.

#Bangladesh snesitivity and specificity
sens <- .824 #corresponding sensitivity
spec <- .810 #corresponding specificity
sens_ll <- .727 #confidence intervals
spec_ll <- .702
sens_ul <- .910
spec_ul <- .906

#north american volunteers
sens_na <- .851 #corresponding sensitivity table S3 (american volunteers)
spec_na <- .947 #corresponding specificity table S3 (american volunteers)

#compute number of samples over 320
overThresh <- data %>% filter(!is.na(vibriocidalMAX_titer)) %>% count(overThresh = (vibriocidalMAX_titer >= thresh))
nResultPos <- overThresh$n[overThresh$overThresh][[1]]
nResultNeg <- overThresh$n[!overThresh$overThresh][[1]]
tot <- nResultPos + nResultNeg

#correct for sensitivity and specificity
corrected <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens, specificity = spec)
corrected_ll_ll <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ll, specificity = spec_ll)
corrected_ll_ul <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ll, specificity = spec_ul)
corrected_ul_ll <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ul, specificity = spec_ll)
corrected_ul_ul <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ul, specificity = spec_ul)

corrected_ll <- min(corrected_ll_ll$realPositives, corrected_ll_ul$realPositives, corrected_ul_ll$realPositives, corrected_ul_ul$realPositives)
corrected_ul <- max(corrected_ll_ll$realPositives, corrected_ll_ul$realPositives, corrected_ul_ll$realPositives, corrected_ul_ul$realPositives)

#north americans. For North Americans Azman et al. don't report a CI.
corrected_na <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_na, specificity = spec_na)


# final AR
attackRateSero <- corrected$realPositives/corrected$totalResults
attackRateSero_ll <- corrected_ll/corrected$totalResults
attackRateSero_ul <- corrected_ul/corrected$totalResults

# north americans
attackRateSero_na <- corrected_na$realPositives/corrected_na$totalResults
```

The naive attack rate using a threshold of 320 on vibriocidal Ogawa is `r 100*nResultPos/corrected$totalResults` %.

Correcting for sensitivity and specificity reported by Azman et al., the attack rate becomes `r 100*attackRateSero` %
using sensitivity and specificity from Bangladesh and `r 100*attackRateSero_na` % using sensitivity and specificity from North American volunteers.


### For under 5s

Under 5

```{r}
#compute number of samples over 320
overThresh <- data %>% filter(!is.na(vibriocidalMAX_titer))  %>% filter(age == "2-4") %>% count(overThresh = (vibriocidalMAX_titer >= thresh))
nResultPos <- overThresh$n[overThresh$overThresh][[1]]
nResultNeg <- overThresh$n[!overThresh$overThresh][[1]]
tot <- nResultPos + nResultNeg

#correct for sensitivity and specificity
corrected <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens, specificity = spec)
corrected_ll_ll <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ll, specificity = spec_ll)
corrected_ll_ul <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ll, specificity = spec_ul)
corrected_ul_ll <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ul, specificity = spec_ll)
corrected_ul_ul <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ul, specificity = spec_ul)

corrected_ll <- min(corrected_ll_ll$realPositives, corrected_ll_ul$realPositives, corrected_ul_ll$realPositives, corrected_ul_ul$realPositives)
corrected_ul <- max(corrected_ll_ll$realPositives, corrected_ll_ul$realPositives, corrected_ul_ll$realPositives, corrected_ul_ul$realPositives)

#north americans
corrected_na <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_na, specificity = spec_na)

# final AR
attackRateSero_u5 <- corrected$realPositives/corrected$totalResults
attackRateSero_u5_ll <- corrected_ll/corrected$totalResults
attackRateSero_u5_ul <- corrected_ul/corrected$totalResults


#north americans
attackRateSero_na_u5 <- corrected_na$realPositives/corrected_na$totalResults
```

### For 5 and over

```{r}
#compute number of samples over 320
overThresh <- data %>% filter(!is.na(vibriocidalMAX_titer))  %>% filter(age == ">=5") %>% count(overThresh = (vibriocidalMAX_titer >= thresh))
nResultPos <- overThresh$n[overThresh$overThresh][[1]]
nResultNeg <- overThresh$n[!overThresh$overThresh][[1]]
tot <- nResultPos + nResultNeg

#correct for sensitivity and specificity
corrected <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens, specificity = spec)
corrected_ll_ll <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ll, specificity = spec_ll)
corrected_ll_ul <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ll, specificity = spec_ul)
corrected_ul_ll <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ul, specificity = spec_ll)
corrected_ul_ul <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_ul, specificity = spec_ul)

corrected_ll <- min(corrected_ll_ll$realPositives, corrected_ll_ul$realPositives, corrected_ul_ll$realPositives, corrected_ul_ul$realPositives)
corrected_ul <- max(corrected_ll_ll$realPositives, corrected_ll_ul$realPositives, corrected_ul_ll$realPositives, corrected_ul_ul$realPositives)

#north americans
corrected_na <- correctForSensSpec(positiveResults = nResultPos, negativeResults = nResultNeg, sensitivity = sens_na, specificity = spec_na)

# final AR
attackRateSero_o5 <- corrected$realPositives/corrected$totalResults
attackRateSero_o5_ll <- corrected_ll/corrected$totalResults
attackRateSero_o5_ul <- corrected_ul/corrected$totalResults

#north americans
attackRateSero_na_o5 <- corrected_na$realPositives/corrected_na$totalResults
```



## Fitting a gaussian mixture model to the data to separate the two distributions:

One way of separating the infected from the non-infected is to assume that the final vibriocidal
log2-titer distribution is the combined log2-distribution for infected and non-infected and that both of them
are gaussian.
The model accounts for the interval censuring that arises from the fact that we don't know where exactly in the
measured titer intervals the exact concentration lies (e.g between a titer value of 1/20 to 1/40).

```{r}
brks <- log2(c(1,10*2^seq(13)))
ggplot(data, aes(vibriocidalMAX_titer_log2)) +
  geom_histogram(aes(y=..density..), breaks = brks, closed = "left") +
  scale_x_continuous(breaks = brks, labels = 2^brks, name = "titer", limits = c(0,NA),
                     sec.axis = sec_axis(~., breaks = brks, name = ("log2(titer)")))
```

Following [Michael Betancourt's analysis](http://mc-stan.org/users/documentation/case-studies/identifying_mixture_models.html) (accessed on 25 October 2023) for the case where we
have sufficient prior information to be able to separate the distributions.

Load rstan package and set options
```{r}
options(mc.cores = parallel::detectCores()) #use in parallel with all cores
rstan_options(auto_write = TRUE) #standard option to aviod recompiling if model code does not change
```

### wide neutral priors

Prepare data and save for stan.
Very wide and non-informative priors, equal for both distributions.
The prior for the mixture parameter theta is set to a beta(1, 1) distribution.

```{r}
intervalBreaks <- log2(10*2^seq(12)) # log2 transform because we fit the normal distributions on a log scale
y <- as.numeric(data$vibriocidalMAX_titer)[!is.na(data$vibriocidalMAX_titer)] #this is the index into the intervals above only!
N <- length(y)
K <- length(intervalBreaks)
prior <- c(6,100,6,100,0,100,1,1) #mu_mu1, sigma_mu1, mu_mu2, sigma_mu2, mu_sigma, sigma_sigma, theta_1, theta_2
nPriorPar <- length(prior)
stan_rdump(c("N", "y", "intervalBreaks", "K", "prior", "nPriorPar"), file=here("src/Stan/mixtureDataStanVibriocidal.R"))
```

Run Stan and analyse posterior
```{r message = FALSE, results = "hide"}
input_data <- read_rdump(here("src/Stan/mixtureDataStanVibriocidal.R"))

fit <- stan(file=here('src/Stan/censuredGaussianMixture2.stan'),
            data=input_data,
            chains=4,
            iter = 5000,
            refresh=1000,
            cores = 4,
            verbose = FALSE,
            control = list(adapt_delta = 0.99), # to avoid divergent transitions
            )
```

```{r}
pars = c("mu","sigma","theta","lp__")
print(fit, pars = pars)
```

```{r}
traceplot(fit, inc_warmup = TRUE, pars = pars)
```

```{r fig.height = 10}
pairs(fit, pars = pars)
```

```{r}
posterior <- extract(fit)

p1 <- ggplot(data, aes(vibriocidalMAX_titer_log2)) +
  geom_histogram(aes(y=..density..), breaks = c(0,intervalBreaks, log2(2*40960)), closed = "left") +
  scale_x_continuous(breaks = c(0,intervalBreaks, log2(2*40960)), labels = 2^c(0,intervalBreaks, log2(2*40960)), name = "titer", limits = c(-1,NA),
                     sec.axis = sec_axis(~., breaks = c(0,intervalBreaks, log2(2*40960)), name = ("log2(titer)")))

for (i in sample(length(posterior$mu[,1]),100)) {
    p1 <- p1 +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = posterior$theta[i], mean = posterior$mu[i,1], sd = posterior$sigma[i,1], log = FALSE), color = "green", alpha = 0.2) +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = 1-posterior$theta[i], mean = posterior$mu[i,2], sd = posterior$sigma[i,2]), color = "blue", alpha = 0.2) +
      stat_function(fun = dnorm_mixture, n = 101, args = list(scale1 = posterior$theta[i], mean1 = posterior$mu[i,1], sd1 = posterior$sigma[i,1], mean2 = posterior$mu[i,2], sd2 = posterior$sigma[i,2]), color = "red", alpha = 0.2)

}

print(p1)
```


In some runs, there is two modes with very close likelihood. This is NOT label switching, it's two real modes.

There is little information about the lower of the two distributions.
Thus we get a posterior with two modes, one where a lot of observations are assumed to be smaller than 1, one with most obs larger than 1.

The mode with the wider lower distribution has a slightly higher likelihood.
We can use stronger priors to favor the mode with the higher likelihood and get STAN to fit better.
We also expect sigma2 to be much lower and of lower uncertainty than sigma 1, so we can try putting different priors on them.

### adjusting priors

```{r}
intervalBreaks <- log2(10*2^seq(12)) # log2 transform because I want to fit the normal distributions on a log scale
y <- as.numeric(data$vibriocidalMAX_titer)[!is.na(data$vibriocidalMAX_titer)] #this is the index into the intervals above only!
N <- length(y)
K <- length(intervalBreaks)
prior <- c(4.5,5,8.5,5,0,20,1,1) #mu_mu1, sigma_mu1, mu_mu2, sigma_mu2, mu_sigma, sigma_sigma, theta_1, theta_2
nPriorPar <- length(prior)
stan_rdump(c("N", "y", "intervalBreaks", "K", "prior", "nPriorPar"), file=here("src/Stan/mixtureDataStanVibriocidal.R"))
```

Run Stan and analyse posterior
```{r message = FALSE, results = "hide"}
input_data <- read_rdump(here("src/Stan/mixtureDataStanVibriocidal.R"))

fit <- stan(file=here('src/Stan/censuredGaussianMixture2.stan'),
            data=input_data,
            chains=4,
            iter = 20000,
            warmup = 16000,
            refresh=1000,
            cores = 4,
            verbose = FALSE,
            control = list(adapt_delta = 0.99), #avoid divergent transitions
          )
```


```{r}
pars = c("mu","sigma","theta","lp__")
print(fit, pars = pars)
```

```{r}
traceplot(fit, inc_warmup = TRUE, pars = pars)
```

```{r fig.height = 10}
pairs(fit, pars = pars)
```

```{r}
posterior <- extract(fit)

p1 <- ggplot(data, aes(vibriocidalMAX_titer_log2)) +
  geom_histogram(aes(y=..density..), breaks = c(0,intervalBreaks, log2(2*40960)), closed = "left") +
  scale_x_continuous(breaks = c(0,intervalBreaks, log2(2*40960)), labels = 2^c(0,intervalBreaks, log2(2*40960)), name = "titer", limits = c(-1,NA),
                     sec.axis = sec_axis(~., breaks = c(0,intervalBreaks, log2(2*40960)), name = ("log2(titer)")))

for (i in sample(length(posterior$mu[,1]),100)) {
    p1 <- p1 +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = posterior$theta[i], mean = posterior$mu[i,1], sd = posterior$sigma[i,1], log = FALSE), color = "green", alpha = 0.2) +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = 1-posterior$theta[i], mean = posterior$mu[i,2], sd = posterior$sigma[i,2]), color = "blue", alpha = 0.2) +
      stat_function(fun = dnorm_mixture, n = 101, args = list(scale1 = posterior$theta[i], mean1 = posterior$mu[i,1], sd1 = posterior$sigma[i,1], mean2 = posterior$mu[i,2], sd2 = posterior$sigma[i,2]), color = "red", alpha = 0.2)

}

print(p1)
```

The attack rate corresponds to 1 - theta:

```{r}
ar_mixture <- 1 - quantile(posterior$theta, c(0.025, .5, .975))
ar_mixture
```



### For >=5 only

```{r}
intervalBreaks <- log2(10*2^seq(12)) # log2 transform because I want to fit the normal distributions on a log scale
y <- as.numeric(data$vibriocidalMAX_titer)[!is.na(data$vibriocidalMAX_titer) & !is.na(data$age) & data$age == ">=5"] #this is the index into the intervals above only!
N <- length(y)
K <- length(intervalBreaks)
prior <- c(4.5,5,8.5,5,3,20,1,1) #mu_mu1, sigma_mu1, mu_mu2, sigma_mu2, mu_sigma, sigma_sigma, theta_1, theta_2
nPriorPar <- length(prior)
stan_rdump(c("N", "y", "intervalBreaks", "K", "prior", "nPriorPar"), file=here("src/Stan/mixtureDataStanVibriocidal.R"))
```

Run Stan and analyse posterior
```{r message = FALSE, results = "hide"}
input_data <- read_rdump(here("src/Stan/mixtureDataStanVibriocidal.R"))

fit <- stan(file=here('src/Stan/censuredGaussianMixture2.stan'),
            data=input_data,
            chains=4,
            iter = 20000,
            warmup = 16000,
            refresh=1000,
            cores = 4,
            verbose = FALSE,
            control = list(adapt_delta = 0.99), #avoid divergent transitions
            )
```


```{r}
pars = c("mu","sigma","theta","lp__")
print(fit, pars = pars)
```

```{r}
traceplot(fit, inc_warmup = TRUE, pars = pars)
```

```{r fig.height = 10}
pairs(fit, pars = pars)
```

```{r}
posterior <- extract(fit)


p1 <- ggplot(data, aes(vibriocidalMAX_titer_log2)) +
  geom_histogram(aes(y=..density..), breaks = c(0,intervalBreaks, log2(2*40960)), closed = "left") +
  scale_x_continuous(breaks = c(0,intervalBreaks, log2(2*40960)), labels = 2^c(0,intervalBreaks, log2(2*40960)), name = "titer", limits = c(-1,NA),
                     sec.axis = sec_axis(~., breaks = c(0,intervalBreaks, log2(2*40960)), name = ("log2(titer)")))

for (i in sample(length(posterior$mu[,1]),100)) {
    p1 <- p1 +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = posterior$theta[i], mean = posterior$mu[i,1], sd = posterior$sigma[i,1], log = FALSE), color = "green", alpha = 0.2) +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = 1-posterior$theta[i], mean = posterior$mu[i,2], sd = posterior$sigma[i,2]), color = "blue", alpha = 0.2) +
      stat_function(fun = dnorm_mixture, n = 101, args = list(scale1 = posterior$theta[i], mean1 = posterior$mu[i,1], sd1 = posterior$sigma[i,1], mean2 = posterior$mu[i,2], sd2 = posterior$sigma[i,2]), color = "red", alpha = 0.2)

}

print(p1)
```

The attack rate corresponds to 1 - theta:

```{r}
ar_mixture_o5 <- 1 - quantile(posterior$theta, c(0.025, .5, .975))
ar_mixture_o5
```

## <5 only

```{r}
intervalBreaks <- log2(10*2^seq(12)) # log2 transform because I want to fit the normal distributions on a log scale
y <- as.numeric(data$vibriocidalMAX_titer)[!is.na(data$vibriocidalMAX_titer) & !is.na(data$age) & data$age == "2-4"] #this is the index into the intervals above only!
N <- length(y)
K <- length(intervalBreaks)
prior <- c(4.5,5,8.5,5,3,20,1,1) #mu_mu1, sigma_mu1, mu_mu2, sigma_mu2, mu_sigma, sigma_sigma, theta_1, theta_2
PriorPar <- length(prior)
stan_rdump(c("N", "y", "intervalBreaks", "K", "prior", "nPriorPar"), file=here("src/Stan/mixtureDataStanVibriocidal.R"))
```

Run Stan and analyse posterior
```{r message = FALSE, results = "hide"}
input_data <- read_rdump(here("src/Stan/mixtureDataStanVibriocidal.R"))

# init <- list(mu1 = 1, mu2 = 8, sigma = c(1,1), theta = 0.5)

fit <- stan(file=here('src/Stan/censuredGaussianMixture2.stan'),
            data=input_data,
            chains=4,
            iter = 20000,
            warmup = 16000,
            refresh=1000,
            cores = 4,
            verbose = FALSE,
            control = list(adapt_delta = 0.999999, max_treedepth = 15), # avoid divergent transitions
            )
```


```{r}
pars = c("mu","sigma","theta","lp__")
print(fit, pars = pars)
```

```{r}
traceplot(fit, inc_warmup = TRUE, pars = pars)
```

```{r fig.height = 10}
pairs(fit, pars = pars)
```

```{r}
posterior <- extract(fit)

p1 <- ggplot(data, aes(vibriocidalMAX_titer_log2)) +
  geom_histogram(aes(y=..density..), breaks = c(0,intervalBreaks, log2(2*40960)), closed = "left") +
  scale_x_continuous(breaks = c(0,intervalBreaks, log2(2*40960)), labels = 2^c(0,intervalBreaks, log2(2*40960)), name = "titer", limits = c(-1,NA),
                     sec.axis = sec_axis(~., breaks = c(0,intervalBreaks, log2(2*40960)), name = ("log2(titer)")))

for (i in sample(length(posterior$mu[,1]),100)) {
    p1 <- p1 +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = posterior$theta[i], mean = posterior$mu[i,1], sd = posterior$sigma[i,1], log = FALSE), color = "green", alpha = 0.2) +
      stat_function(fun = dnorm_rescale, n = 101, args = list(scale = 1-posterior$theta[i], mean = posterior$mu[i,2], sd = posterior$sigma[i,2]), color = "blue", alpha = 0.2) +
      stat_function(fun = dnorm_mixture, n = 101, args = list(scale1 = posterior$theta[i], mean1 = posterior$mu[i,1], sd1 = posterior$sigma[i,1], mean2 = posterior$mu[i,2], sd2 = posterior$sigma[i,2]), color = "red", alpha = 0.2)

}

print(p1)
```

The attack rate corresponds to 1 - theta:

```{r}
ar_mixture_u5 <- 1 - quantile(posterior$theta, c(0.025, .5, .975))
ar_mixture_u5
```


# Results table & Figure

```{r}
attack_rate <- 
  tribble(
    ~type, ~ar_type, ~age_cat, ~central, ~low, ~high,
    "clinical surveillance",                                               "clinical", "all", attackRate_clinical, NA, NA,
    "clinical surveillance",                                               "clinical", "<5",  attackRate_clinical_u5, NA, NA,
    "clinical surveillance",                                               "clinical", "2-4 years",  attackRate_clinical_2to4, NA, NA,
    "clinical surveillance",                                               "clinical", ">=5 years", attackRate_clinical_o5, NA, NA,
    "clinical surveillance",                                               "clinical", "All>=2 years", attackRate_clinical_o2, NA, NA,
    "self reported cholera",                                               "clinical", "All>=2 years", ar_self_rep$chol, NA, NA,
    "self reported watery diarrhea",                                       "clinical", "All>=2 years", ar_self_rep$wd, NA, NA,
    "self reported cholera",                                               "clinical", "2-4 years", ar_self_rep$chol_2to4, NA, NA,
    "self reported watery diarrhea",                                       "clinical", "2-4 years", ar_self_rep$wd_2to4, NA, NA,
    "self reported cholera",                                               "clinical", ">=5 years", ar_self_rep$chol_geq5, NA, NA,
    "self reported watery diarrhea",                                       "clinical", ">=5 years", ar_self_rep$wd_geq5, NA, NA,
    "vibriocidal threshold\n(test preformance Bangladesh)",                "infection",  "All>=2 years", attackRateSero, attackRateSero_ll, attackRateSero_ul,
    "vibriocidal threshold\n(test preformance Bangladesh)",                "infection",  "2-4 years", attackRateSero_u5, attackRateSero_u5_ll, attackRateSero_u5_ul,
    "vibriocidal threshold\n(test preformance Bangladesh)",                "infection",  ">=5 years", attackRateSero_o5, attackRateSero_o5_ll, attackRateSero_o5_ul,
    "vibriocidal threshold\n(test preformance North America)",             "infection",  "All>=2 years", attackRateSero_na, NA, NA,
    "vibriocidal threshold\n(test preformance North America)",             "infection",  "2-4 years", attackRateSero_na_u5, NA, NA,
    "vibriocidal threshold\n(test preformance North America)",             "infection",  ">=5 years", attackRateSero_na_o5, NA, NA,
    "vibriocidal gaussian mixture",                                        "infection", "All>=2 years", ar_mixture[["50%"]], ar_mixture[["97.5%"]], ar_mixture[["2.5%"]],
    "vibriocidal gaussian mixture",                                        "infection", "2-4 years", ar_mixture_u5[["50%"]], ar_mixture_u5[["97.5%"]], ar_mixture_u5[["2.5%"]],
    "vibriocidal gaussian mixture",                                        "infection", ">=5 years", ar_mixture_o5[["50%"]], ar_mixture_o5[["97.5%"]], ar_mixture_o5[["2.5%"]]
  )
```

```{r}
attack_rate %>%
  mutate(ci = paste0(round(10*100*central)/10, "% [", round(10*100*low)/10, " - ", round(10*100*high)/10, "]")) %>%
  rmarkdown::paged_table()
```


```{r}
attack_rate %>%
  filter(type %in% c("clinical surveillance", "self reported cholera", "self reported watery diarrhea", "vibriocidal threshold\n(test preformance Bangladesh)", "vibriocidal threshold\n(test preformance North America)", "vibriocidal gaussian mixture"),
      age_cat %in% c("All>=2 years", "2-4 years", ">=5 years")
   ) %>%
  mutate(ci = paste0(round(10*100*central)/10, "% [", round(10*100*low)/10, " - ", round(10*100*high)/10, "]")) %>%
  select(type, age_cat, ci) %>%
  pivot_wider(names_from = age_cat, values_from = ci) %>%
  rmarkdown::paged_table()
```

```{r fig.width = 6, fig.height = 8}
cols <- c("#708090", "#f08080")
types <- c("clinical surveillance", "self reported cholera", "self reported watery diarrhea", "vibriocidal threshold\n(test preformance Bangladesh)", "vibriocidal threshold\n(test preformance North America)", "vibriocidal gaussian mixture")

attack_rate %>%
  filter(
    type %in% types,
    age_cat %in% c("All>=2 years", "2-4 years", ">=5 years"),
    ) %>%
  mutate(
    type = factor(type, levels = types),
    age_cat = factor(age_cat, levels = c("All>=2 years", "2-4 years", ">=5 years"))
  ) %>%
  ggplot(aes(y = type, x = 100*central, xmin = 100*low, xmax = 100*high, col = ar_type)) +
  geom_pointrange(position = position_dodge2(width = .4), orientation = "y", size = .75) +
  theme_bw() +
  facet_grid(rows = vars(age_cat)) +
  lims(x = c(0, 100)) +
  labs(
    x = "Attack rate (%)",
    y = "Estimate"
  ) +
  scale_y_discrete(limits = rev(types)) +
  scale_color_manual(values = cols) +
  theme(legend.position = "none")
```


# R Environment

Output environment and package versions used for reproducibility:

```{r}
sessionInfo()

stan_version()
```
