---
title: "Logistic regression - association between serology and self-reported cholera"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: true
    df_print: paged
---

```{r echo = FALSE, warning = FALSE, include = FALSE}
# Setup Environment
knitr::opts_chunk$set(
  echo = TRUE,
  include = TRUE,
  eval = TRUE,
  warning = TRUE,
  message = FALSE,
  cache = F,
  prompt = F,
  comment = NA,
  fig.width = 7,
  fig.height = 4
  )

# Use english date format even with french or german locale.
Sys.setenv(LANGUAGE="en")
```

# Goal

Derive the association between serology and self-reported cholera using the data from  [Jackson et al. (2013)](https://doi.org/10.4269%2Fajtmh.13-0208).

# Libraries

```{r}
# install.packages(c("here", "dplyr", "tidyr", "magrittr", "readr", "gtsummary"))

library(here)
library(dplyr)
library(tidyr)
library(magrittr)
library(gtsummary)
```


# Data

Note that this analysis requires individual level data, which isn’t included in this repository to protect privacy of the survey participants. Interested researchers are invited to contact the authors.

```{r}
data <- readr::read_csv(here("data/serology_individual.csv"))

#assign 1 to titers not detectable
#change wrong label for highest dilution
levels(data$vibriocidalMAX_titer)[levels(data$vibriocidalMAX_titer)=="40560"] <- "40960"
levels(data$vibriocidalMAX_titer)[levels(data$vibriocidalMAX_titer)=="1.414"] <- "1"

#transform titers to log2 scale
data %<>%
  mutate(
    vibriocidalMAX_titer_log2 = log2(as.integer(as.character(vibriocidalMAX_titer)))
  )
```


# Data summary

```{r}
gtsummary::tbl_summary(
    data %>% select(
      sex,
      age,
      location,
      vibriocidalMAX_titer_log2,
      cholera,
      watdiar
    )
  )
```


# Outcome: self reported cholera

## Univariate regression

```{r}
d_model <- data %>%
    transmute(
          sex,
          age,
          location,
          vibriocidalMAX_titer_log2,
          cholera
        ) %>%
    replace_na(
        list(
            sex = "missing",
            location = "missing"
        )
    )

out_lr_uni <- d_model %>%
  gtsummary::tbl_uvregression(
        method = glm,
        y = cholera,
        method.args = list(family = binomial),
        exponentiate = TRUE
  ) %>%
    gtsummary::bold_labels() %>%
    gtsummary::italicize_levels()
```


## Data for multivariate model

Keep only variables with p < 0.2 in the univariate model. 

```{r}
sig_vars <- out_lr_uni$table_body %>% filter(p.value < 0.2) %>% distinct(variable)

d_model %<>%
  select(cholera, sig_vars$variable) %>%
  na.omit()
```

## Multivariate regression without location

Location has significant and non-significant levels. Test if including it improves the fit.

```{r}
formula <- cholera ~ age + vibriocidalMAX_titer_log2

m_lr_multi_nl <- glm(formula, data = d_model, family = binomial(link = "logit"))

summary(m_lr_multi_nl)
```

## Multivariate regression with location

```{r}
formula <- cholera ~ age + vibriocidalMAX_titer_log2 + location

m_lr_multi <- glm(formula, data = d_model, family = binomial(link = "logit"))

summary(m_lr_multi)
```

Compare with and without location:

```{r}
anova(m_lr_multi_nl, m_lr_multi, test = "Chisq")
```

Adding location does improve fit.

## Final output

Univariate and multivariate models in one table:

```{r}
out_lr_step <- m_lr_multi %>%
  gtsummary::tbl_regression(
        exponentiate = TRUE
  ) %>%
    gtsummary::bold_labels() %>%
    gtsummary::italicize_levels()

gtsummary::tbl_merge(list(out_lr_uni, out_lr_step), tab_spanner = c("**Univariate**", "**Multivariate**"))
```



# Outcome: self reported watery diarrhoea

## Univariate regression

```{r}
d_model <- data %>%
    transmute(
          sex,
          age,
          location,
          vibriocidalMAX_titer_log2,
          watdiar
        ) %>%
    replace_na(
        list(
            sex = "missing",
            location = "missing"
        )
    )

out_lr_uni <- d_model %>%
  gtsummary::tbl_uvregression(
        method = glm,
        y = watdiar,
        method.args = list(family = binomial),
        exponentiate = TRUE
  ) %>%
    gtsummary::bold_labels() %>%
    gtsummary::italicize_levels()
```


## Data for multivariate model

Keep only variables with p < 0.2 in the univariate model. 

```{r}
sig_vars <- out_lr_uni$table_body %>% filter(p.value < 0.2) %>% distinct(variable)

d_model %<>%
  select(watdiar, sig_vars$variable) %>%
  na.omit()
```

## Multivariate regression without location

Location has significant and non-significant levels. Test if including it improves the fit.

```{r}
formula <- watdiar ~ age + vibriocidalMAX_titer_log2

m_lr_multi_nl <- glm(formula, data = d_model, family = binomial(link = "logit"))

summary(m_lr_multi_nl)
```

## Multivariate regression with location

```{r}
formula <- watdiar ~ age + vibriocidalMAX_titer_log2 + location

m_lr_multi <- glm(formula, data = d_model, family = binomial(link = "logit"))

summary(m_lr_multi)
```

Compare with and without location:

```{r}
anova(m_lr_multi_nl, m_lr_multi, test = "Chisq")
```

Adding location does improve fit.

## Final output

Univariate and multivariate models in one table:

```{r}
out_lr_step <- m_lr_multi %>%
  gtsummary::tbl_regression(
        exponentiate = TRUE
  ) %>%
    gtsummary::bold_labels() %>%
    gtsummary::italicize_levels()

gtsummary::tbl_merge(list(out_lr_uni, out_lr_step), tab_spanner = c("**Univariate**", "**Multivariate**"))
```


# R Environment

Output environment and package versions used for reproducibility:

```{r}
sessionInfo()
```